\section{Scope, claim status, and reconstruction rule}

The preceding ITSM manuscript combined an active-superfluid ontology, a
Born--Infeld interpolation, an algebraic galactic force, an open-system source
and multiple cosmological claims.  Subsequent equation audits and numerical
gates showed that these ingredients do not yet follow from one action.  The
purpose of this alpha is therefore architectural: it states what the model is,
which results are presently supportable, and what must still be calculated.

Four labels are used throughout:

\begin{description}
\item[Derived.] A checked result that follows from declared assumptions within
its stated domain.
\item[Conditional.] A result that also requires a named postulate, matching
hypothesis, background choice or symmetry limit.
\item[Open.] A well-defined derivation or numerical test remains incomplete.
\item[Rejected.] The stated derivation or implementation failed and is not live
support for the framework.
\end{description}

These labels apply to derivations rather than to broad themes.  For example,
topology-dependent Casimir stress is derived for the declared periodic scalar
calculation, while a universal $13/12$ expansion ratio from cycle counting is
rejected.  A rejected shortcut does not reject $\Tthree$ topology; likewise, a
successful local effective action does not establish a relativistic or
cosmological completion.

\subsection{Conventions}

We use metric signature $(-,+,+,+)$ and natural units $c=\hbar=1$ except
where factors are restored explicitly, as in the Casimir lattice energy.  The
reduced Planck mass and gravitational coupling satisfy
\begin{equation}
 \mpl^{-2}=8\pi G,
 \qquad
 \kappa=8\pi G.
 \label{eq:conventions}
\end{equation}
These conventions make the covariant $Y^{3/2}$ coefficient and the static
cubic-gradient coefficient directly comparable.

\begin{table}[ht]
\centering
\small
\caption{High-level status at the start of the core recovery.}
\begin{tabular}{p{0.36\textwidth}p{0.18\textwidth}p{0.34\textwidth}}
\toprule
Component & Status & Boundary of the result \\
\midrule
Finite-density plenum & Conditional & Physical postulate; microscopic action open \\
Compact $\Tthree$ topology & Conditional & Adopted boundary condition; scale open \\
$Y^{3/2}$ square-root branch & Conditional & Local, quasi-static, high symmetry \\
Present Born--Infeld origin & Rejected & Does not give the required IR response \\
$\cproj=2/3$ local vertex & Conditional & Geometric motivation; matching open \\
Rectangular-$\Tthree$ Casimir tensor & Derived & Periodic free scalar with stated moduli \\
Persistent $13/12$ from free-scalar Casimir stress & Rejected & CBR-001 finds only transient crossings \\
Persistent $13/12$ from a derived driven ITSM sector & Open & Requires TOP/WAK/reservoir closure and CBR-002 \\
Screening, PPN and lensing & Open & Relativistic completion absent \\
Realistic disk field solution & Open & Periodic nonlinear PDE required \\
Driven reservoir anisotropy & Open & Constitutive conversion not derived \\
\bottomrule
\end{tabular}
\end{table}

The reconstruction rule is that no observational claim may have a stronger
status than its upstream field sector.  Historical calculations remain useful
as diagnostics and provenance, but they are not promoted into the core merely
because their numerical values are suggestive.
