\section{Core fields and layered architecture}

\subsection{Metric, condensate, and finite-density frame}

The gravitational variable is the spacetime metric $g_{\mu\nu}$.  The
candidate plenum order parameter is complex,
\begin{equation}
 \Phi=\frac{\rho}{\sqrt{2}}e^{i\Theta},
 \label{eq:complex-order}
\end{equation}
so that amplitude fluctuations, phase excitations, zeros of the order
parameter and winding sectors can be represented within one microscopic
language.  This is important because a smooth single real potential is locally
irrotational, whereas the historical ITSM ontology relies on circulation,
vortices and toroidal and poloidal currents.

A unit timelike vector $U^\mu$ denotes the rest frame of the finite-density
state,
\begin{equation}
 U^\mu U_\mu=-1,
 \qquad
 h^{\mu\nu}=g^{\mu\nu}+U^\mu U^\nu.
 \label{eq:projector}
\end{equation}
Whether $U^\mu$ is derived from $\Theta$, from a foliation field, or from a
coarse-grained fluid description is an open choice that must be settled by the
relativistic gate.

\subsection{Emergent phonon and matter}

The infrared phonon is denoted $\psi$.  It is an excitation of the
finite-density phase, not automatically the same variable as a fundamental
vacuum scalar at all scales.  Matter fields are collected in $\Psi_m$.  The
matter--phonon interaction must be derived once, then used consistently both
in the phonon source and in the matter stress-energy exchange.

\subsection{Topology, reservoir, and non-equilibrium sectors}

Spatial slices are assigned compact $\Tthree$ boundary conditions with lengths
$L_i$.  Their ratios are shape moduli.  The global topology controls allowed
modes and winding; it does not directly fix every local Wilson coefficient.

An explicit reservoir sector $R$ permits the observable matter--plenum system
to be open while the complete covariant system remains conserved.  A wake or
memory variable $W$ and a shape sector are permitted research sectors, not
assumed completed degrees of freedom.

The schematic inventory is
\begin{equation}
 S_{\rm complete}=
 S_{\rm EH}+S_{\Phi}+S_T+S_m+S_R
 +S_{\rm int}+S_W,
 \label{eq:inventory}
\end{equation}
where $S_W$ is included only if a causal and stable wake description is found.
At low energy, $S_\Phi+S_{\rm int}$ may be matched to an infrared phonon action
$S_{\rm IR}$.  One must not add the UV and IR descriptions as independent
propagating copies in the same regime; they are Wilsonian descriptions linked
by matching.

\subsection{Architecture versus prediction}

Equation~\eqref{eq:inventory} is a sector map, not a declaration that each term
has been derived.  The architecture recovers the original active and
rotational character of ITSM while making the missing links explicit.  In
particular,
\begin{equation}
 \text{energy throughput}
 \not\Rightarrow
 \text{anisotropic pressure}
\end{equation}
without a shape-dependent conversion law, and
\begin{equation}
 \text{compact topology}
 \not\Rightarrow
 \cproj=\frac23 \quad\text{or}\quad \frac{H_t}{H_p}=\frac{13}{12}
\end{equation}
without separate matching and backreaction calculations.
