\section{Ultraviolet--infrared hierarchy}

\subsection{Candidate analytic condensate}

A representative microscopic starting point is
\begin{equation}
 S_\Phi=\int \dd^4x\sqrt{-g}\left[
 -\nabla_\mu\Phi^*\nabla^\mu\Phi
 -m^2|\Phi|^2
 -\frac{\lambda_4}{2}|\Phi|^4
 -\frac{\lambda_6}{3\Lambda^2}|\Phi|^6
 +\cdots\right].
 \label{eq:candidate-uv}
\end{equation}
Equation~\eqref{eq:candidate-uv} is a matching candidate, not the canonical
ITSM action.  Its branch structure, finite-density state, stability and
relativistic characteristics must be evaluated before adoption.

The sextic term is of interest because a three-body density interaction can
generate the power required by the infrared gate.  In a Thomas--Fermi
description, consider
\begin{equation}
 \mathcal L(n,X)=nX-\frac{\lambda_3}{3}n^3.
 \label{eq:aux-density}
\end{equation}
The algebraic density equation gives
\begin{equation}
 n=\sqrt{\frac{X}{\lambda_3}},
\end{equation}
and substitution yields
\begin{equation}
 P(X)=\frac{2}{3\sqrt{\lambda_3}}X^{3/2}.
 \label{eq:sextic-px}
\end{equation}
This establishes a plausible mechanism for the power law, not the complete
relativistic ITSM matching.  Gradient corrections, sign choices, the domain of
$X$, background stability and normalization remain open.

\subsection{Conditional infrared action}

Define the dimensionless spatial invariant
\begin{equation}
 Y=\frac{h^{\mu\nu}\nabla_\mu\psi\nabla_\nu\psi}{a_0^2}.
 \label{eq:Y}
\end{equation}
The conditional low-acceleration operator is
\begin{equation}
 \mathcal L_{\psi,\rm IR}
 =-\frac{2C_{\rm IR}}{3}\mpl^2a_0^2Y^{3/2}
 +\mathcal O(Y^2,\nabla^2/\Lambda_{\rm IR}^2).
 \label{eq:ir-action}
\end{equation}
Here $C_{\rm IR}$ is a Wilson coefficient in a chosen field normalization.  It
cannot be identified with a local observable coupling until the matter vertex
has been normalized in the same convention.

\subsection{Why the preceding Born--Infeld claim is not retained}

The preceding manuscript used an analytic kinetic expression whose derivative
approaches a nonzero constant at small kinetic invariant.  A sourced field
equation is therefore linear at leading order in that limit.  The desired
square-root response instead requires a kinetic derivative proportional to the
field-gradient magnitude, as supplied by Eq.~\eqref{eq:ir-action}.  The
Born--Infeld expression may be investigated for a separate UV or regulating
role, but it is not the present derivation of the galactic branch.

\subsection{UVIR-001 pass criteria}

A successful match must demonstrate:
\begin{enumerate}
\item a stable finite-density background;
\item a heavy amplitude mode that can be integrated out;
\item the sign and branch of Eq.~\eqref{eq:ir-action};
\item positive compressibility and Hamiltonian in the domain of use;
\item a controlled sound cone and strong-coupling scale;
\item the coefficient map to $G$, $a_0$ and the normalized matter vertex.
\end{enumerate}
Until these checks pass, Eq.~\eqref{eq:ir-action} is a conditional infrared
completion rather than a derived consequence of Eq.~\eqref{eq:candidate-uv}.
