\section{Wake, circulation, and non-equilibrium hypotheses}

The original ITSM described baryonic matter as exciting a torsional or
acoustic wake in a rotational medium.  The static equation
Eq.~\eqref{eq:periodic-p3} has no independent memory: for specified boundary
data it tracks the instantaneous source.  A persistent, advected or detached
wake therefore requires additional dynamics.

A schematic relaxation equation is
\begin{equation}
 \tau_W U^\mu\nabla_\mu W+W
 =\mathcal F[\psi,\rho_b,r,Q_{\rm syn}],
 \label{eq:wake-schematic}
\end{equation}
or, if wave propagation is essential, a hyperbolic field equation with a
second time derivative.  Equation~\eqref{eq:wake-schematic} is not adopted as a
law.  It identifies the quantities that WAK-001 must derive or bound.

A viable wake sector must satisfy all of the following:
\begin{enumerate}
\item a well-posed initial-value problem;
\item causal characteristic speeds in the physical frame;
\item a positive or bounded energy accounting;
\item a controlled static limit reproducing the IR field equation;
\item explicit momentum exchange with matter and the reservoir;
\item decay or transport laws rather than indefinite unsupported memory.
\end{enumerate}

Similarly, circulation is represented naturally by the phase in
Eq.~\eqref{eq:complex-order},
\begin{equation}
 \oint \nabla_i\Theta\,\dd x^i=2\pi n,
 \qquad n\in\mathbb Z,
\end{equation}
where zeros of $\rho$ permit defect cores.  The relationship between these
microscopic winding sectors and a global toroidal current is the VOR-001 gate.
No black-hole, lensing or galactic coefficient follows from the winding number
without further dynamics.

For driven anisotropy one may organize the bookkeeping as
\begin{equation}
 \Pi_{\rm total}=\Pi_{\rm Cas}+\Pi_{\rm shape}
 +\Pi_{\rm driven}+\Pi_{\rm memory}.
 \label{eq:pi-decomposition}
\end{equation}
Only $\Pi_{\rm Cas}$ has been calculated in the present programme.  The other
terms are labels for research sectors.  Writing Eq.~\eqref{eq:pi-decomposition}
does not establish that they exist or have the required sign and magnitude.
