\section{Cosmological architecture and current limits}

The homogeneous condensate, reservoir and topology moduli must define one
fiducial cosmological background before CMB, growth, age or distance results
are calculated.  Earlier analyses used several dataset-conditioned values of
the syntropic exponent or effective equation of state and sometimes mixed
them across diagnostics.  Those outputs remain provenance and cannot be
treated as one canonical cosmology.

At background level, a conservative sector decomposition has the schematic
form
\begin{align}
 \dot\rho_P+3H(\rho_P+p_P)+\mathcal W_\Pi &= Q,\\
 \dot\rho_R+3H(\rho_R+p_R) &= -Q,
 \label{eq:background-exchange}
\end{align}
where $\mathcal W_\Pi$ denotes the work performed by anisotropic stress in the
chosen geometry.  Its exact coefficient follows from the covariant stress
tensor, not from notation.  The total continuity equation follows by summing
the sectors.

No predetermined law such as $Q\propto(1+z)^{-3}$ is adopted here.  In
particular, $(1+z)^{-3}=a^3$ grows toward late time and cannot simultaneously
be described as an early-universe enhancement.  COS-001 must derive the source
and fit one background with an explicit parameter inventory.

\subsection{What CBR-001 changes}

CBR-001 does not remove topology-dependent cosmological stress.  It shows that
the tested passive free-field stress redshifts away and cannot support a
persistent $13/12$ directional expansion ratio.  Therefore
\begin{itemize}
\item $H_0=72.97\,\mathrm{km\,s^{-1}\,Mpc^{-1}}$ is not a parameter-free ITSM
prediction;
\item assigning Planck and SH0ES to opposite directional extrema is not a
derived observation model;
\item any future directional expansion must be passed through actual survey
window functions;
\item a driven anisotropic state requires a derived source beyond free Casimir
stress.
\end{itemize}

\subsection{Perturbations}

PERT-001 must vary the same total action and exchange currents used for the
background.  Effective CAMB or CLASS mappings are useful diagnostics, but they
do not replace the derivation of scalar, vector and tensor perturbations,
initial conditions, stability and observable transfer functions.  Until that
gate closes, CMB-peak and $S_8$ values are diagnostic calculations rather than
canonical predictions.
