\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage[colorlinks=true,allcolors=blue]{hyperref}

\newcommand{\mpl}{M_{\mathrm P}}
\newcommand{\cproj}{C_{\mathrm{proj}}}
\newcommand{\cmp}{Q_{\mathrm{mp}}}
\newcommand{\csyn}{Q_{\mathrm{syn}}}
\newcommand{\dd}{\mathrm d}
\newcommand{\Tthree}{T^3}

\title{The Integrated Toroidal--Syntropic Model:\\
Core Architecture, Conditional Weak-Field Limit, and Open Research Gates}
\author{Brendon Boyd\\Independent Researcher, Western Australia}
\date{Version 12.0-alpha --- Core Recovery Draft --- \today}

\begin{document}

\maketitle

\begin{center}
\fbox{\parbox{0.92\textwidth}{\textbf{Research-stage status.}
This document reconstructs a coherent ITSM field hierarchy and records both
successful and failed research gates. It is not a completed alternative to
$\Lambda$CDM and does not claim that the open matching, screening, lensing,
disk, perturbation, or driven-anisotropy problems have been solved.}}
\end{center}

\begin{abstract}
The Integrated Toroidal--Syntropic Model (ITSM) is reconstructed here as a
layered effective framework rather than a single scalar interpolation formula.
The recovered architecture contains a finite-density complex condensate, an
emergent infrared phonon sector, distinct matter--plenum and
reservoir--plenum exchange currents, compact three-torus boundary conditions,
and explicitly open wake, circulation, shape-modulus, screening and
relativistic sectors.  A cubic-gradient phonon action proportional to
$Y^{3/2}$ conditionally yields the desired square-root acceleration in a
quasi-static high-symmetry limit; it is not derived from the analytic
Born--Infeld term used in the preceding manuscript.  On compact $\Tthree$ the
local source must be compensated and realistic disks require a nonlinear
periodic boundary-value solution.  Separately, the validated CBR-001 lattice
and biaxial backreaction calculations establish topology-dependent directional
Casimir pressure but find that $H_t/H_p=13/12$ is only a tunable transient in
the tested free-field model, not a plateau or attractor.  The present work
therefore defines a conservative theory architecture, a strict claim-status
vocabulary and a dependency-ordered programme for microscopic matching,
screening, lensing, disks, wakes and driven anisotropic stress.
\end{abstract}

\tableofcontents

\input{sections/01_scope_status}
\input{sections/02_core_architecture}
\input{sections/03_uv_ir_hierarchy}
\input{sections/04_conservation_exchange}
\input{sections/05_weak_field}
\input{sections/06_topology_casimir}
\input{sections/07_wake_nonequilibrium}
\input{sections/08_cosmological_status}
\input{sections/09_open_gates}
\input{sections/10_falsifiability}

\section*{Conclusion}

The recovery preserves the defining ITSM idea---a driven, finite-density,
rotational medium with compact topology and sector exchange---while removing
the unsupported shortcuts that had come to carry that idea.  The local
$Y^{3/2}$ weak-field branch and the rectangular-$\Tthree$ Casimir tensor are
concrete advances.  Their microscopic connection to a condensate, their
relativistic completion and their consequences for realistic systems remain
research gates.  Future predictive claims must be downstream of those gates.

\end{document}
